Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-11-27 revise/revise.tex
Make revision compile.

\question In the introduction, the authors mention that: ``Two key advantages
offered by the fluid-fuelled MSR are improved fuel utilization and no radiation
damage constraint on attainable fuel burn-up. Together, these attributes result
in significantly reduced core fissile inventories and spent nuclear fuel mass''
However, the fissile inventory is not necessarily related to fuel burnup and
fuel utilization. It can actually be very high in chloride-based fast-spectrum
MSRs.
\begin{solution}
        We have edited this sentence to read ``Together these attributes can
        result in significantly reduced spent fuel mass and, in the case of
        fluoride salt reactors in particular, reduced core fissile
        inventories'', recognizing that you're quite correct for the chloride salts.
\end{solution}

@@

Manuele Aufiero, Antonio Cammi, Olivier Geoffroy, Mario Losa, Lelio Luzzi, Marco
E. Ricotti, Hervé Rouch, Development of an OpenFOAM model for the Molten Salt
Fast Reactor transient analysis, Chemical Engineering Science, Volume 111, 2014,
Pages 390-401, ISSN 0009-2509, http://dx.doi.org/10.1016/j.ces.2014.03.003.
(http://www.sciencedirect.com/science/article/pii/S0009250914001146)

Carlo Fiorina, Danny Lathouwers, Manuele Aufiero, Antonio Cammi, Claudia
Guerrieri, Jan Leen Kloosterman, Lelio Luzzi, Marco Enrico Ricotti, Modelling
and analysis of the MSFR transient behaviour, Annals of Nuclear Energy, Volume
64, 2014, Pages 485-498, ISSN 0306-4549,
http://dx.doi.org/10.1016/j.anucene.2013.08.003.
(http://www.sciencedirect.com/science/article/pii/S0306454913004118)

Matteo Zanetti, Antonio Cammi, Carlo Fiorina, Lelio Luzzi, A Geometric
Multiscale modelling approach to the analysis of MSR plant dynamics, Progress in
Nuclear Energy, Volume 83, 2015, Pages 82-98, ISSN 0149-1970,
http://dx.doi.org/10.1016/j.pnucene.2015.02.014.

A. Laureau, D. Heuer, E. Merle-Lucotte, P.R. Rubiolo, M. Allibert, M. Aufiero,
Transient coupled calculations of the Molten Salt Fast Reactor using the
Transient Fission Matrix approach, Nuclear Engineering and Design, Volume 316,
2017, Pages 112-124, ISSN 0029-5493,
http://dx.doi.org/10.1016/j.nucengdes.2017.02.022.

@@

        In a similar vein, Kophazi et al. used iterative coupling between
        in-house
        three-dimensional neutronic and one-dimensional heat conduction models
        DALTON
        and THERM to analyze normal MSRE operation as well as
        channel-blocking-incident transients \cite{kophazi_development_2009}.
        The
        Kophazi model added entrance effects of heat transfer coefficients as
        well as
        thermal coupling between fuel channels through moderator heat
        conduction. More
        recently, Cammi et al. performed a 2D-axisymmetric single-channel
        analysis of
        the MSBR using the commercial finite element package COMSOL
        Multiphysics
        \cite{cammi_multi-physics_2011}. That work directly solved the fuel
        salt
        velocity field, used heterogeneous group constants in fuel and
        moderator
        regions, and employed the COMSOL software package intrinsically
        designed
        for coupled multi-physics simulation.  Fiorina, Lathouwers, and their
        colleagues conducted a benchmarking exercise
        \cite{fiorina_modelling_2013} in
        which this Politecnico di Milano approach was expanded to a
        multi-channel model
        of the MSFR and compared to code from the University of Delft
        \cite{de_zwaan_static_2007,van_der_linden_2012} based on the approach
        in
        \cite{kophazi_development_2009}. These models showed good agreement for
        multiple accident transients. Meanwhile, leveraging lessons learned
        from these
        efforts has resulted in a multiscale approach from Zanetti et al.
        \cite{zanetti_geometric_2015} successfully combines high and low
        geometric
        fidelity for graphite-moderated MSRs.

@@

        In this work, $Q_g = \gamma V_{\text{core}}^{-1} \int_{\text{core}}Q_f
        \diff V$, where $\gamma$ is a factor representing heat dissipation by
        gamma and neutron irradiation in the moderator, called the graphite to
        fuel power density ratio. Robertson's \cite{robertson_conceptual_1971}
        original \gls{MSBR} analysis included a calculation of $\gamma$.
        \cite{krepel_dyn3d-msr_2007,zhang_development_2009,cammi_multi-physics_2011}
        all calculated gamma heating through such a factor. We follow
        \cite{cammi_multi-physics_2011} and set $\gamma=0.0144$. Notably, the
        Moltres physics kernel for radiation heating can be calculate the volume
        average fission heat in a variety of means: whole-core and local
        averages are possible. Whole-core averages are employed for simplicity
        and in accordance with prior literature.  Knowledge of fision heat
        rates, however, require knowledge of neutron fluxes, implying a need for
        group constants.
\end{solution}

@@

       The base 2D mesh has a characteristic size of 10 cm in the axial
       direction and .4 cm in the radial direction to capture the variation from moving
       between fuel and moderator subdomains. To determine whether results were
       converged, a mesh convergence study was conducted with up to three levels of isotropic
       refinement (e.g. each element was split in half in both axial and radial directions,
       resulting in four new elements). The metric used to assess convergence was the integrated fast group
       flux. The result of the mesh convergence study is shown in
       \cref{fig:mesh_convergence}. From refinement level 2 to 3 the metric changes by
       only 1\%. We regard the results as sufficiently converged at level 2 for the
       purposes of our study; consequently, the results reported in the following
       section correspond to a radial element dimension of .1 cm and an axial element
       dimension of 2.5 cm.
\end{solution}

@@

        Although the qualitative agreement is decent, there are discrepancies
        between the \gls{MSRE} and Moltres calculations. As outlined above there
        is an 11 K offset between \gls{MSRE} and Moltres calculations for the
        fuel temperature in the hottest channel. The peak graphite temperature
        is around 14 K larger in the Moltres calculation. Fast fluxes are larger
        in the \gls{MSRE} calculation by roughly 20\%. These are not
        insignificant differences. However, given the differing nature of the
        two models--no axial conduction in the \gls{MSRE} model, 2-group vs
        32-group neutronics, exclusion of the control rod thimbles in the
        Moltres calculation--we believe this variation in quantitative behavior
        is acceptable for the purpose of this work, which is to introduce
        Moltres as a simulation tool. One additional feature that could
        influence Moltres results is inclusion of precursor decay heat; this is
        under active development. Trustworthy verification of Moltres results,
        such as spatial flux and temperature profiles, will require more
        detailed experimental measurements than those available in the
        \gls{ORNL} technical reports, which contain only macroscopic data such
        as steady-state power and heat-exchanger temperature
        drops. Additionally, detailed comparison with other modern modeling
        efforts such as that conducted in
        \cite{aufiero_development_2014,laureau_transient_2017} is desirable. In
        that vein, a collaboration is under way with the primary author of
        \cite{aufiero_develoment_2014} to try and reproduce calculation results
        from OpenFOAM. We hope to publish the results of this comparison in a
        future work.
\end{solution}

@@

\end{questions}
\bibliographystyle{unsrt}
\bibliography{Moltres}
  \end{document}

>>>>

\question In the introduction, the authors mention that: ``Two key advantages
offered by the fluid-fuelled MSR are improved fuel utilization and no radiation
damage constraint on attainable fuel burn-up. Together, these attributes result
in significantly reduced core fissile inventories and spent nuclear fuel mass''
However, the fissile inventory is not necessarily related to fuel burnup and
fuel utilization. It can actually be very high in chloride-based fast-spectrum
MSRs.
\begin{solution}
        This passage has been modified to read:

        The key advantages of MSRs generally pertain to improved fuel
        utilization and reactor safety. In contrast to legacy reactors, only
        moderator fast neutron damage and fuel chemistry evolution limit
        burnup. A clever configuration of moderator as in
        \cite{engel_conceptual_1980} can enable reactor operation without
        opening the vessel for thirty or more years. Further, several fission
        products selectively precipitate onto nickel surfaces in fluoride salt,
        as documented in \cite{engel_conceptual_1980}, thus reducing unwanted
        neutron absorption. Lastly, the epithermal spectrum of
        graphite-moderated salt reactors incinerates plutonium more efficiently,
        thus reducing long-lived transuranic waste production
        \cite{engel_conceptual_1980}.  The sum of these characteristics implies
        the MSR uniqely ameliorates spent fuel burden whilst extending
        nuclear fuel resources. To top all these benefits off, xenon transients
        become moot in MSRs due to its insolubility in salt, thus
        narrowing transient analysis focus to thermalhydraulic concerns.

@@

Manuele Aufiero, Antonio Cammi, Olivier Geoffroy, Mario Losa, Lelio Luzzi, Marco
E. Ricotti, Hervé Rouch, Development of an OpenFOAM model for the Molten Salt
Fast Reactor transient analysis, Chemical Engineering Science, Volume 111, 2014,
Pages 390-401, ISSN 0009-2509

Carlo Fiorina, Danny Lathouwers, Manuele Aufiero, Antonio Cammi, Claudia
Guerrieri, Jan Leen Kloosterman, Lelio Luzzi, Marco Enrico Ricotti, Modelling
and analysis of the MSFR transient behaviour, Annals of Nuclear Energy, Volume
64, 2014, Pages 485-498, ISSN 0306-4549

Matteo Zanetti, Antonio Cammi, Carlo Fiorina, Lelio Luzzi, A Geometric
Multiscale modelling approach to the analysis of MSR plant dynamics, Progress in
Nuclear Energy, Volume 83, 2015, Pages 82-98, ISSN 0149-1970

A. Laureau, D. Heuer, E. Merle-Lucotte, P.R. Rubiolo, M. Allibert, M. Aufiero,
Transient coupled calculations of the Molten Salt Fast Reactor using the
Transient Fission Matrix approach, Nuclear Engineering and Design, Volume 316,
2017, Pages 112-124, ISSN 0029-5493

@@

        In a similar vein, Kophazi et al. used iterative coupling between
        in-house
        three-dimensional neutronic and one-dimensional heat conduction models
        DALTON
        and THERM to analyze normal MSRE operation as well as
        channel-blocking-incident transients \cite{kophazi_development_2009}.
        The
        Kophazi model added entrance effects of heat transfer coefficients as
        well as
        thermal coupling between fuel channels through moderator heat
        conduction. More
        recently, Cammi et al. performed a 2D-axisymmetric single-channel
        analysis of
        the MSBR using the commercial finite element package COMSOL
        Multiphysics
        \cite{cammi_multi-physics_2011}. That work directly solved the fuel
        salt
        velocity field, used heterogeneous group constants in fuel and
        moderator
        regions, and employed the COMSOL software package intrinsically
        designed
        for coupled multi-physics simulation.  Fiorina, Lathouwers, and their
        colleagues conducted a benchmarking exercise
        \cite{fiorina_modelling_2014} in
        which this Politecnico di Milano approach was expanded to a
        multi-channel model
        of the MSFR and compared to code from the University of Delft
        \cite{de_zwaan_static_2007,van_der_linden_coupled_2012} based on the approach
        in
        \cite{kophazi_development_2009}. These models showed good agreement for
        multiple accident transients. Meanwhile, leveraging lessons learned
        from these
        efforts has resulted in a multiscale approach from Zanetti et al.
        \cite{zanetti_geometric_2015} successfully combines high and low
        geometric
        fidelity for graphite-moderated MSRs.

@@

        In this work, $Q_g = \gamma V_{\text{core}}^{-1} \int_{\text{core}}Q_f dV$,
        where $\gamma$ is a factor representing heat dissipation by
        gamma and neutron irradiation in the moderator, called the graphite to
        fuel power density ratio. Robertson's \cite{robertson_conceptual_1971}
        original MSBR analysis included a calculation of $\gamma$.
        \cite{krepel_dyn3d-msr_2007,zhang_development_2009,cammi_multi-physics_2011}
        all calculated gamma heating through such a factor. We follow
        \cite{cammi_multi-physics_2011} and set $\gamma=0.0144$. Notably, the
        Moltres physics kernel for radiation heating can be calculate the volume
        average fission heat in a variety of means: whole-core and local
        averages are possible. Whole-core averages are employed for simplicity
        and in accordance with prior literature.  Knowledge of fision heat
        rates, however, require knowledge of neutron fluxes, implying a need for
        group constants.
\end{solution}

@@

       The base 2D mesh has a characteristic size of 10 cm in the axial
       direction and .4 cm in the radial direction to capture the variation from moving
       between fuel and moderator subdomains. To determine whether results were
       converged, a mesh convergence study was conducted with up to three levels of isotropic
       refinement (e.g. each element was split in half in both axial and radial directions,
       resulting in four new elements). The metric used to assess convergence was the integrated fast group
       flux. The result of the mesh convergence study is shown in
       figure []. From refinement level 2 to 3 the metric changes by
       only 1\%. We regard the results as sufficiently converged at level 2 for the
       purposes of our study; consequently, the results reported in the following
       section correspond to a radial element dimension of .1 cm and an axial element
       dimension of 2.5 cm.
\end{solution}

@@

        Although the qualitative agreement is decent, there are discrepancies
        between the MSRE and Moltres calculations. As outlined above there
        is an 11 K offset between MSRE and Moltres calculations for the
        fuel temperature in the hottest channel. The peak graphite temperature
        is around 14 K larger in the Moltres calculation. Fast fluxes are larger
        in the MSRE calculation by roughly 20\%. These are not
        insignificant differences. However, given the differing nature of the
        two models--no axial conduction in the MSRE model, 2-group vs
        32-group neutronics, exclusion of the control rod thimbles in the
        Moltres calculation--we believe this variation in quantitative behavior
        is acceptable for the purpose of this work, which is to introduce
        Moltres as a simulation tool. One additional feature that could
        influence Moltres results is inclusion of precursor decay heat; this is
        under active development. Trustworthy verification of Moltres results,
        such as spatial flux and temperature profiles, will require more
        detailed experimental measurements than those available in the
        ORNL technical reports, which contain only macroscopic data such
        as steady-state power and heat-exchanger temperature
        drops. Additionally, detailed comparison with other modern modeling
        efforts such as that conducted in
        \cite{aufiero_development_2014,laureau_transient_2017} is desirable. In
        that vein, a collaboration is under way with the primary author of
        \cite{aufiero_development_2014} to try and reproduce calculation results
        from OpenFOAM. We hope to publish the results of this comparison in a
        future work.
\end{solution}

@@

\end{questions}
\bibliographystyle{unsrt}
\bibliography{revise}
  \end{document}
